---
title: "Fieldworker deployment Mozambique"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, echo=F, warning=F, results=F}
library(tidyverse)
library(magrittr)
library(ggmap)
# library(igraph)
library(geosphere)
library(data.table)
library(DT)
library(TSP)

SOURCE_DIR <- "../SourceData"
DERIVED_DIR <- "../DerivedData"

SOURCE_HOUSEHOLDS <- paste( SOURCE_DIR, "2021-09-23 - Household Data.csv", sep="/")
DERIVED_FULLY_CONNECTED <- paste( DERIVED_DIR, "2021-09-23 - FullyConnected.csv", sep="/")
DERIVED_HOUSEHOLDS_CORE <- paste( DERIVED_DIR, "2021-09-26 - Households_core.gdf", sep="/")
```

Calculate the fieldworkers required for questionnaires in all households in Mozambique. This is done by

* finding a route that passes all households in a cluster exactly once,
* assuming an average travel speed, and
* estimating visit duration from size of household (using [a linear model](https://github.com/databrew/bohemia/tree/master/scripts/fullcensus/report)).

Clusters are then classified by criticality, i.e. which clusters require more work than a field worker (work unit) can do between visits.

NOTE: this Rmd does not rely on igraph functionality

# Setup
```{r echo=F}
df.raw.households <- read.csv( SOURCE_HOUSEHOLDS, stringsAsFactors = F)

head(df.raw.households)
```
Quick look at the data
```{r}
summary( df.raw.households )
```

## Visualize

```{r}
moz <- c(left = 35.29, bottom = -18.2, right = 36.57, top = -17.52)

map.moz <- get_map(location=moz, source="stamen", maptype="watercolor", crop=FALSE)

ggmap( map.moz ) +
  geom_point( data=df.raw.households %>% filter(core_or_buffer=="buffer"), aes(lng, lat, color=core_or_buffer), size=0.8 ) +
  geom_point( data=df.raw.households %>% filter(core_or_buffer=="core"), aes(lng, lat, color=core_or_buffer), size=0.1, alpha=0.3 ) +
  scale_color_manual( values=c("core"="orange", "buffer"="darkred")) +
  labs( title="Households Mozambique",
        x="Longitude",
        y="Latitude",
        color="Role in cluster")
```
# Estimate time effort per cluster
## Work load per household
We assume 20 mins base time for each visit and an additional 15" for each member of the household according to Joe's linear model. Values are in [h].

```{r}
df.workload <- df.raw.households %>%
                        mutate( visit_duration = 20/60 + n_people*15/60 ) %>% # 20" base, 15" per person ...
                        select( hh_id, visit_duration, lng, lat, cluster )

df.workload %<>% 
  group_by( cluster ) %>% 
  summarise( cluster_visit_time = sum(visit_duration),
             lng = mean(lng),
             lat = mean(lat),
             n = n()) %>% 
  ungroup()
```

## Tour all households by cluster
Here we look for a route visiting all households in a cluster exactly once. Given an average speed we can take a guess at how long this may take.

The following helper function creates a symmetric adjacency matrix that can be processed by the TSP solver. The solver returns a route in [m].

```{r}
get_tsp_distance <- function( cluster_no, verbose=0 ) {
  if( verbose==1 )
    sprintf( "Cluster #%d", cluster_no )
  df <- df.raw.households %>% 
    filter( cluster == cluster_no )

  nodes <- unique(df$hh_id)
  order <- length(nodes)
  edges <- data.frame( matrix(0L, nrow=order, ncol=order))
  colnames(edges) <- nodes
  rownames(edges) <- nodes
  edges$start <- nodes
  
  trips <- edges %>% 
    pivot_longer(!start, names_to = "dest", values_to = "dist")
  
  # add geospatial data for both nodes
  trips <- merge( trips, df.raw.households %>% select(hh_id, lat, lng), by.x="start", by.y="hh_id", all.x=T)
  trips <- merge( trips,  df.raw.households %>% select(hh_id, lat, lng), by.x="dest", by.y="hh_id", all.x=T)
  
  # calculate distances
  trips$dist <- distHaversine(trips[,4:5], trips[,6:7])
  
  edges <-
    trips %>% 
    select( start, dest, dist ) %>% 
    pivot_wider( names_from = dest, values_from = dist) %>% 
    arrange(start) %>% 
    select(sort(current_vars())) %>% # sort by col names
    select(-start)
  
  colnames(edges) <-c() # if we don't do this we get a 'not symmetric matrix' error ...
  
  distances <- data.matrix(edges)

  tsp <- TSP(distances)
  tour <- solve_TSP(tsp)
  
  return(attr(tour, "tour_length"))
}
```

For cluster #1:
```{r}
print( get_tsp_distance(cluster_no = 1) )
```

Do this for every cluster. Assume an average speed of 3 km/h and 30" to pack up and leave. This could be adapted to cluster location if desired.

```{r}
travel_speed <- 3 # kph
df.workload %<>%
  group_by(cluster) %>% 
  mutate( tour = get_tsp_distance(cluster)) %>% 
  mutate( cluster_travel_time = n*0.5 + (tour/1000)/travel_speed)

df.workload
```
# Results 
Everything in place to summarise. A field worker's maximum work capacity is put at 24 workdays @7h (8h with 1h break).

## Distribution of work load
```{r}
field_worker_capacity <- (8-1)*24

df.workload %<>% 
  mutate( total_time = cluster_visit_time + cluster_travel_time )

ggplot( df.workload, aes(total_time)) +
  geom_histogram(fill="lightblue") +
  geom_vline( xintercept = field_worker_capacity, linetype="dashed") +
  annotate( "text", x=50+field_worker_capacity, y=50, label="24 d") +
  labs( title = "Time required per cluster",
        x = "Total time [h]",
        y = "Number of clusters") +
  theme_minimal()
```

## Total estimates
Total workload for site in [h]

```{r}
site_work_load <- sum(df.workload$total_time)
site_work_load
```

Given field worker capacity, calculate number of field workers required.

Note: this is an upper bound because here every single household participates. In reality, this will be higher because of backtracking to re-visit households where nobody was in. Additionally, while routing would need to be optimized, the real route taken by field workers will also be determined by 

* personal preference (familiarity with terrain)
* environmental factors like rainfall requiring re-routing

```{r}
site_work_load / field_worker_capacity
```

## Visualization
Criticality categories:
- >95% of total workload = red
- 80-95% of total workload = yellow
- <80% of total workload = green

```{r}
max_time <- max(df.workload$total_time)
df.workload %<>%
  mutate( node_weight = total_time / max_time,
          criticality = case_when(
            total_time > field_worker_capacity*.95 ~ "red",
            total_time < field_worker_capacity*.8 ~ "green",
            TRUE ~ "yellow"
          ))

ggmap( map.moz ) +
  geom_point( data=df.workload, aes(lng, lat, size=node_weight, color=criticality), alpha=0.5 ) +
  scale_color_manual(values=c("red" = "darkred", "yellow"="sienna2", "green" = "darkgreen")) +
  labs( title="Workload by Cluster (Mozambique)",
        subtitle="Relative working hours estimate",
        caption="24 8h workdays, travel at 3kph, visits 20 min +15 min per person",
        x="Longitude",
        y="Latitude",
        color="Criticality") +
  guides(size=F)
```

## Interactive table
```{r}
datatable(df.workload %>% select(-lng, -lat, -node_weight))
```

